---
title: "Bernauer_Andrew_report2"
author: "Andrew Bernauer"
date: "February 12, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)
```
#Report 1


##Introduction 
For the semester long project I will be attempting to predict the known response variable ethereum price in terms of the predictors: block size, network hash rate growth rate, transactions, day of the week, and month. The project is regression based and not a classification task. Therefore the project will fall under the umbrella of supervised machine learning.    



$$ether price = \beta_{o} + \beta_{1} \times blocksize + \beta_{2} \times networkhashgrowthrate + \beta_{3}  \times transactions + \beta_{4} \times day + \beta_{5} \times month + \epsilon $$

Ethereum is a decentralized, open source, block chain technology, featuring smart contracts. The crypto currency that fuels the Ethereum block chain is Ether. 



###Data  
```{r data carpentry, cache=TRUE}


library(scales)
library(ggplot2)
library(lubridate)
library(purrr)
library(dplyr)
library(readr)
library(tibble)
library(errorist)
library(broom)
library(caret)

#substitute in the path on your machine to files 
ethereum_transaction_history <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumTransactionHistory.csv")

ethereum_block_size <-read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumBlockSizeHistory.csv"
)

ethereum_network_hash_rate <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumNetworkHashRateGrowthRate.csv")

ether_price <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EtherPriceHistory(USD).csv")

# creating variable for day  
day_of_week <- 
  lubridate::mdy(ether_price$`Date(UTC)`) %>%
  wday( ,label=TRUE)  %>%
  sort()

# creating variable for month
month_of_year <-
  lubridate::mdy(ether_price$`Date(UTC)`) %>%
  month( ,label=TRUE) %>%
  sort()

signedlog10 <- function(x) {
  ifelse(abs(x) <= 1, 0, sign(x)*log10(abs(x)))
}

log_price <- signedlog10(ether_price[[3]])

log_history <- signedlog10(ethereum_transaction_history[[3]])

log_block_size <- signedlog10(ethereum_block_size[[3]])

log_hash_rate <- signedlog10(ethereum_network_hash_rate[[3]])

summary(log_price)

summary(log_history)

# construct list of variables to coerce into a tibble
l_ether <- list( price = ether_price[[3]],
                 transaction_history = as.integer(ethereum_transaction_history[[3]]),
                 block_size = as.integer(ethereum_block_size[[3]]),
                 hash_rate = ethereum_network_hash_rate[[3]], 
                 day = day_of_week, 
                 month = month_of_year,
                 date_utc = mdy(ethereum_network_hash_rate$`Date(UTC)`),
                 unix_time_stamp = ethereum_network_hash_rate$UnixTimeStamp,
                 log_10_price = log_price,
                 log_10_trans_history = log_history,
                 log_10_block_size = log_block_size,
                 log_10_hash_rate = log_hash_rate)

ether_df <- as_tibble(l_ether)


xs <-function(x){(x-mean(x))/(2*sd(x))}

```



####Plots
```{r price_vs_time, warning=FALSE, cache=TRUE}
 
price_history <- ggplot(ether_df) +
  aes(date_utc, price) +
  geom_line(color="purple") + 
  xlab("Year")+
  ggtitle("Ethereum Price USD over time") + 
  scale_y_log10() 
  
  
  

 price_history
```


The price doesn't seem move at all significantly until 2016 so I decided to use a log transformation to the y axis. This removed a large amount the visual noise and made the price change overtime more evident. Besides that ethereum price starts to rise from around a dollar price to hovering around the ten dollars between 2016 and 2017. Peaks at a value of 1000 dollars in 2018.   





```{r block_size, cache=TRUE}
block_history <- ggplot(ether_df) +
  aes(date_utc, block_size) +
  geom_line(color ="purple") +
  scale_y_log10() +
  xlab("Year") +
  ylab("Block Size") +
  ggtitle("Ethereum Block Size Over Time")
  

block_history
```

Block size rises exponentially begining in 2017. Prior to this it fluctuates up and down between 1000 and 3000 which might be pointing to autocorrelation. In this same time period it peaks violently at 10000. Reaching a maximum of over 30000 in 2018. Another log tranformation was used in this plot.  



```{r network hashrate, cache=TRUE}
 hashrate_history <- ggplot(ether_df) + 
  aes(date_utc, hash_rate) +
  geom_line(color="purple") +
  scale_y_log10() +
  xlab("Year") +
  ylab("Network Hash rate Growth rate") +
  ggtitle("Ethereum Network Hash rate growth over time")

 hashrate_history
```

Network hashrate growth blows up exponentially and seems to follow that trend overtime. 

```{r transaction history, warning=FALSE, cache=TRUE}
 transaction_history <- ggplot(ether_df) + 
  aes(date_utc, transaction_history) +
  geom_line(color="purple") +
  xlab("Year") +
  ylab("Transactions") +
  scale_y_log10() +
  ggtitle("Ethereum Transactions over time")

 transaction_history
```

Transactions on the the Ethereum network are growing significantly overtime and more people are embracing it's use.





```{r bonus plot from eda, warning= FALSE, cache=TRUE}
 price_plot <- ggplot(ether_df) +
   aes(transaction_history, price) +
   geom_point(colour="purple") + 
   facet_wrap(~month) +
   geom_smooth(colour="yellow", alpha = 0.25) 

 price_plot +
   scale_x_log10("Transactions") +
   scale_y_log10("Ethereum Price USD($)") +
   ggtitle("Ethereum Price USD($) vs Transactions facetted by month") 
 
```

There appears to be some clustering of Ether price as transactions increases on the block chain increase isolating by month.



```{r}
p <- ggplot(ether_df) +
 aes(log_10_hash_rate) + geom_histogram(bins = 34, fill = "purple") +
  ggtitle(label = "Histogram of the distribution of hash rate log 10 scale") +
  xlab("Distribution of log 10 transactions")
  
  
p
```

Many of the $X_{1}, X_{2}, \dots, X_{n}$ or predictor variables in the dataset follow skewed distribution and feature multiple peaks. This could be pointing to the presence of sub-populations within the underlying data. 


```{r hist block_size,}
ggplot(ether_df) +
  aes(log_block_size) + geom_histogram(bins = 34, fill = 'purple') +
  xlab("Distribution of Ethereum BlockSize log 10 scale") +
  ylab("Count") +
  ggtitle("Histogram of Ethereum block size")

```

Likewise, blocksize has this similar property this sort of distribution doesn't suit itself well standard machine learning algorithms given the multiple peaks. Most of the variables follow gamma distributions in the cartesian coordinate scale. I declined to illustrate those graphs.


#####Summary Code

```{r tidy_summary_code}

#summary stats for ethereum price  
ether_df %>% summarise(mean_price = mean(price, na.rm = TRUE), median_price = median(price), sd_price = sd(price), iqr_price = IQR(price), n = n(), mad_price = mad(price), min_price = min(price), max_price = max(price))

#summary stats for ethereum transaction history
ether_df %>% summarise(mean_transaction_history = mean(transaction_history), median_transaction_history = median(transaction_history), sd_transaction_history = sd(transaction_history), iqr_transaction_history = IQR(transaction_history), n = n(), mad_transaction_history = mad(transaction_history), min_transaction_history = min(transaction_history), max_transaction_history = max(transaction_history))

#summary stats for ethereum block size
ether_df %>% summarise(mean_block_size = mean(block_size), median_block_size = median(block_size), sd_block_size = sd(block_size), iqr_block_size = IQR(block_size), n = n(), mad_block_size = mad(block_size), min_block_size = min(block_size), max_block_size = max(block_size))

#summary stats for hash rate
ether_df %>% summarise(mean_hash_rate = mean(hash_rate), median_hash_rate = median(hash_rate), sd_hash_rate= sd(hash_rate), iqr_hash_rate = IQR(hash_rate), n = n(), mad_hash_rate = mad(hash_rate), min_hash_rate = min(hash_rate), max_hash_rate = max(hash_rate))




```


#Report 2





##Regressions


```{r regressions,}
#running five regressions 
 
poly_reg <- lm(price ~ poly(transaction_history, 5) + poly(block_size, 5) + poly(hash_rate, 5), ether_df )


reg_obj <- lm(log(price) ~ log(block_size) + log(transaction_history) + log(hash_rate), data = ether_df,  subset = price > 0)

reg_obj_2 <- lm(log(price) ~ log(transaction_history) + log(hash_rate), data = ether_df, subset = price > 0)


reg_obj_3 <- lm(price ~ I(block_size)^2 + sqrt(transaction_history) + hash_rate, data = ether_df)

mols <- lm(price ~ block_size + transaction_history + hash_rate, data = ether_df)

log_ols <- lm(log(price) ~ log(transaction_history), data = ether_df, subset = price > 0)


 

```



###Regression Diagnostics 

The following code block returns the adjusted $R^{2}$ statistics for the five regressions I ran.



```{r}
glance_tidy_poly <- glance(poly_reg)

glance_reg_obj <- glance(reg_obj)

glance_reg_obj_2 <- glance(reg_obj_2)

glance_reg_obj_3 <- glance(reg_obj_3)

glance_mols <- glance(mols)

glance_log_ols <- glance(log_ols)


glance_tidy_poly$adj.r.squared

glance_reg_obj$adj.r.squared

glance_reg_obj_2$adj.r.squared

glance_reg_obj_3$adj.r.squared

glance_mols$adj.r.squared

glance_log_ols$adj.r.squared

```
The log-log model with all the continuous predictors included had the highest adjusted $R^{2} = 0.9363884 $ of all the regressions run. A log-log model can be interpreted as a the percentage change in Y by a one percentage change in X. Accordingly, if X increase by 1 we expect a $\beta_{1}$ percent change in Y. 


The model in this case is $\ln(ether price) = \beta_{0} + $





```{r, cache=TRUE}
tidy_reg_obj <- reg_obj %>% 
   tidy()

tidy_reg_obj
```





```{r augment, cache=TRUE}
augmented_reg_obj <- reg_obj %>%
             augment()

some_plot <- ggplot(data = augmented_reg_obj, aes(log.price., .hat)) +
  geom_point(color = "purple", size = 1) +
  labs(x="log(price)", y="y hat") +
  ggtitle("Log price vs Y hat") + 
  scale_y_log10()
  

some_plot
```
The preceding plot points to the non-linear nature of the data set. It may improve modeling results to include a polynomial term in future models.





```{r, warning=FALSE}
tidy_conf_it <- tidy(reg_obj, conf.int = TRUE)

confint_plot <- ggplot(tidy_conf_it, aes(term, estimate, color=term)) +
  geom_point() +
  geom_errorbar(aes(ymin=estimate - 1.96*std.error, ymax=estimate + 1.96*std.error)) 
  

confint_plot +
  ggtitle("Coefficients with standard errors") +
  scale_y_discrete()
  

```




```{r proper scale, warning=FALSE}

confint_plot + 
  scale_y_log10() +
  ggtitle("Coefficients with standard errors log-log scale") + 
  scale_x_discrete()

```





























```{r}
some_plot_2 <- ggplot(data = augmented_reg_obj, aes(.resid)) + 
  geom_histogram(binwidth = .3, fill = "purple") + 
  labs(x="Residuals", y="Frequency") +
  ggtitle("Distribution of Residuals")
  
  

some_plot_2
```
The majority of the residuals are distributed between -1 and 1 peaking at zero. This distribution is similar to the laplace distribution. 



```{r, warning= FALSE}
some_plot_3 <- ggplot(augmented_reg_obj, aes(log.price., .resid)) +
  geom_point(color = "purple", size = 1) +
  geom_hline(color = "yellow", yintercept = 0) 

some_plot_3 + 
  labs(x="Log Ether Price", y="Residuals")




```







##KNN Regression







```{r knn regression, eval=FALSE, cache=TRUE}
library(caret)

set.seed(20)

train_index <- createDataPartition(ether_df$price, p = .8, list = FALSE, times = 1)

head(train_index)

ether_df_train <- ether_df[train_index, ]

ether_df_test <- ether_df[-train_index, ]

tr_control <- trainControl(method = "repeatedcv",
             number = 10,
             repeats = 3
             )
knn_model <-  train(price ~., data = ether_df,
                    method = "knn",
                    preProcess = c("center", "scale"), 
                    trControl = tr_control, 
                    metric = "RMSE",
                    tuneLength = 10
                    )

knn_model

knn_model$results

y_hat <- predict(knn_model, newdata = ether_df)

predicted <- predict(knn_model, ether_df)

plot(ether_df$price, predicted)
sqrt(sum((predicted - ether_df$price) ^ 2) / length(ether_df)) 

```


```{r 2) }
library(glmnet)
library(caret)


set.seed(385)


train_df <- ether_df %>% sample_frac(0.5)
test_df <- ether_df %>% setdiff(train_df)

train_df <- train_df[, -9]

test_df <- test_df[, -9]

covariates_train <- model.matrix(price~. , data = train_df)[, c(-1, -9)]

covariates_test <- model.matrix(price~. , data=test_df)[, c(-1, -9)]
response_train <- as.numeric(unlist(train_df[1]))
response_test <- as.numeric((unlist(test_df[1])))

grid <- 10^seq(10, -2, length.out = 100)

ridge_model <- glmnet(covariates_train, response_train, alpha = 0, lambda = grid, family ="gaussian")

plot.glmnet(ridge_model, xvar = "lambda")

lasso_model <- glmnet(covariates_test, response_train, alpha = 1, lambda = grid, family = "gaussian")

ls(lasso_model)
plot.glmnet(lasso_model, xvar = "lambda")




```



```{r 2a)}
set.seed(385)

cv_lasso <- cv.glmnet(covariates_train, train_df$price, alpha=1)


plot(cv_lasso)

set.seed(385)
cv_ridge <- cv.glmnet(covariates_train, train_df$price, alpha=0)

plot(cv_ridge)

best_lambda_ridge <- cv_ridge$lambda.1se

best_lambda_lasso <- cv_lasso$lambda.1se

print(best_lambda_ridge)


ridge_coefficients <- coef.glmnet(cv_ridge, newx = covariates_train,  s = best_lambda_ridge)



lasso_coef <- coef.glmnet(cv_lasso, newx = covariates_train, s = best_lambda_lasso)

print(lasso_coef)

print(ridge_coefficients)


```


```{r}
lasso_test_pred = predict.glmnet(lasso_model, newx = covariates_test, s=best_lambda_lasso)

ridge_test_pred = predict.glmnet(ridge_model, newx = covariates_test, s=best_lambda_ridge)
mean((response_test - lasso_test_pred) ^ 2)

mean((response_test - ridge_test_pred) ^ 2)
```


```{r}
library(tree)
set.seed(385)

ether_tree <- tree(price~., data = train_df)
summary(ether_tree)
plot(ether_tree)
text(ether_tree, pretty = 0)
cv_tree <- cv.tree(ether_tree)
plot(cv_tree$size, cv_tree$dev, type = 'b')

pruned_tree_df <- prune.tree(ether_tree, best = 5)
plot(pruned_tree_df)
text(pruned_tree_df, pretty = 0)


```


