---
title: "Bernauer_Andrew_report2"
author: "Andrew Bernauer"
date: "February 12, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Report 1


##Introduction 
For the semester long project I will be attempting to predict the known response variable ethereum price in terms of the predictors: block size, network hash rate growth rate, transactions, day of the week, and month. The project is regression based and not a classification task. Therefore the project will fall under the umbrella of supervised machine learning.    



$$ethereum price = \beta_{o} + \beta_{1} \times blocksize + \beta_{2} \times networkhashgrowthrate + \beta_{3}  \times transactions + \beta_{4} \times day + \beta_{5} \times month + \epsilon $$

Ethereum is a decentralized, open source, block chain technology, featuring smart contracts. The crypto currency that fuels the Ethereum block chain is Ether. Block size refers to the size 



###Data  
```{r data carpentry}


library(scales)
library(ggplot2)
library(lubridate)
library(purrr)
library(dplyr)
library(readr)
library(tibble)
library(errorist)

#substitute in the path on your machine to files 
ethereum_transaction_history <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumTransactionHistory.csv")

ethereum_block_size <-read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumBlockSizeHistory.csv"
)

ethereum_network_hash_rate <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EthereumNetworkHashRateGrowthRate.csv")

ether_price <- read_csv("C:\\Users\\andre\\Documents\\ECON_490_ML\\ML_report 1\\ethereum-historical-data\\EtherPriceHistory(USD).csv")

# creating variable for day  
day_of_week <- 
  lubridate::mdy(ether_price$`Date(UTC)`) %>%
  wday( ,label=TRUE) %>%
  sort()

# creating variable for month
month_of_year <-
  lubridate::mdy(ether_price$`Date(UTC)`) %>%
  month( ,label=TRUE) %>%
  sort()

# construct list of variables to coerce into a tibble
l_ether <- list( price = ether_price$Value,
                 transaction_history = as.integer(ethereum_transaction_history$Value),
                 block_size = as.integer(ethereum_block_size$Value),
                 hash_rate = ethereum_network_hash_rate$Value, 
                 day = day_of_week, 
                 month = month_of_year,
                 date_utc = mdy(ethereum_network_hash_rate$`Date(UTC)`),
                 unix_time_stamp = ethereum_network_hash_rate$UnixTimeStamp)

ether_df <- as_tibble(l_ether)

head(ether_df)

tail(ether_df)

summary(ether_df)

```



####Plots
```{r price_vs_time, warning=FALSE}
 
price_history <- ggplot(ether_df) +
  aes(date_utc, price) +
  geom_line(color="purple") + 
  xlab("Year")+
  ggtitle("Ethereum Price USD over time") + 
  scale_y_log10() 
  
  
  

 price_history
```


The price doesn't seem move at all significantly until 2016 so I decided to use a log transformation to the y axis. This removed a large amount the visual noise and made the price change overtime more evident. Besides that ethereum price starts to rise from around a dollar price to hovering around the ten dollars between 2016 and 2017. Peaks at a value of 1000 dollars in 2018.   





```{r block_size}
block_history <- ggplot(ether_df) +
  aes(date_utc, block_size) +
  geom_line(color ="purple") +
  scale_y_log10() +
  xlab("Year") +
  ylab("Block Size") +
  ggtitle("Ethereum Block Size Over Time")
  

block_history
```

Block size rises exponentially begining in 2017. Prior to this it fluctuates up and down between 1000 and 3000 which might be pointing to autocorrelation. In this same time period it peaks violently at 10000. Reaching a maximum of over 30000 in 2018. Another log tranformation was used in this plot.  



```{r network hashrate}
 hashrate_history <- ggplot(ether_df) + 
  aes(date_utc, hash_rate) +
  geom_line(color="purple") +
  scale_y_log10() +
  xlab("Year") +
  ylab("Network Hash rate Growth rate") +
  ggtitle("Ethereum Network Hash rate growth over time")

 hashrate_history
```

Network hashrate growth blows up exponentially and seems to follow that trend overtime. 

```{r transaction history, warning=FALSE}
 transaction_history <- ggplot(ether_df) + 
  aes(date_utc, transaction_history) +
  geom_line(color="purple") +
  xlab("Year") +
  ylab("Transactions") +
  scale_y_log10() +
  ggtitle("Ethereum Transactions over time")

 transaction_history
```

Transactions on the Ethereum network are growing very quickly and people are embracing the technology it is based on.





```{r bonus plot from eda, warning= FALSE}
 price_plot <- ggplot(ether_df) +
   aes(transaction_history, price) +
   geom_point(colour="purple") + 
   facet_wrap(~month) +
   geom_smooth(colour="yellow", alpha = 0.25) 

 price_plot +
   scale_x_log10("Transactions") +
   scale_y_log10("Ethereum Price USD($)") +
   ggtitle("Ethereum Price USD($) vs Transactions facetted by month") 
 
```

Ether Price appears to fall into clusters as transactions increase and you isolate for month of the year. 




#####Summary Code

```{r tidy_summary_code, eval=FALSE}

#summary stats for ethereum price  
ether_df %>% summarise(mean_price = mean(price, na.rm = TRUE), median_price = median(price), sd_price = sd(price), iqr_price = IQR(price), n = n(), mad_price = mad(price), min_price = min(price), max_price = max(price))

#summary stats for ethereum transaction history
ether_df %>% summarise(mean_transaction_history = mean(transaction_history), median_transaction_history = median(transaction_history), sd_transaction_history = sd(transaction_history), iqr_transaction_history = IQR(transaction_history), n = n(), mad_transaction_history = mad(transaction_history), min_transaction_history = min(transaction_history), max_transaction_history = max(transaction_history))

#summary stats for ethereum block size
ether_df %>% summarise(mean_block_size = mean(block_size), median_block_size = median(block_size), sd_block_size = sd(block_size), iqr_block_size = IQR(block_size), n = n(), mad_block_size = mad(block_size), min_block_size = min(block_size), max_block_size = max(block_size))

#summary stats for hash rate
ether_df %>% summarise(mean_hash_rate = mean(hash_rate), median_hash_rate = median(hash_rate), sd_hash_rate= sd(hash_rate), iqr_hash_rate = IQR(hash_rate), n = n(), mad_hash_rate = mad(hash_rate), min_hash_rate = min(hash_rate), max_hash_rate = max(hash_rate))




```

#####Conclusion

Ethereum data is quite complex on the surface level before the next report I should complete some more research on it's block chain technology. As I am more familiar with Bitcoin and it could lead to new insights on the data. 



#Report 2: Regression Modeling Ether Price





#Regressions


```{r regressions,}
#running five regressions 
 
poly_reg <- lm(price ~ poly(transaction_history, 5) + poly(block_size, 5) + poly(hash_rate, 5), ether_df )


reg_obj <- lm(log(price) ~ log(block_size) + log(transaction_history) + log(hash_rate), data = ether_df,  subset = price > 0)

reg_obj_2 <- lm(log(price) ~ log(transaction_history) + log(hash_rate), data = ether_df, subset = price > 0)


reg_obj_3 <- lm(price ~ I(block_size)^2 + sqrt(transaction_history) + hash_rate, data = ether_df)

mols <- lm(price ~ block_size + transaction_history + hash_rate, data = ether_df)

log_ols <- lm(log(price) ~ log(transaction_history), data = ether_df, subset = price > 0)


 

```



##Regression Diagnostics 

The following code chunk returns tidyed summary for the best regression I ran.
```{r tidy diagnostics}
library(broom)


tidy_reg_obj <- reg_obj %>%
  tidy()


tidy_reg_obj
glance(reg_obj)
```
All of the coefficients return p values significant at the .10 cut off level while log block size is close. 




```{r}
glance_tidy_poly <- glance(poly_reg)

glance_reg_obj <- glance(reg_obj)

glance_reg_obj_2 <- glance(reg_obj_2)

glance_reg_obj_3 <- glance(reg_obj_3)

glance_mols <- glance(mols)

glance_log_ols <- glance(log_ols)


glance_tidy_poly$adj.r.squared

glance_reg_obj$adj.r.squared

glance_reg_obj_2$adj.r.squared

glance_reg_obj_3$adj.r.squared

glance_mols$adj.r.squared

glance_log_ols$adj.r.squared

```
The log-log model with all the continuous predictors included had the highest adjusted $R^2 $ of all the regressions run. A model excluding the log block size term comes in a close second. Another, potential option is the last model with just log transaction history as a predictor. 

Predictors performing the best presented the following equation $\hat{y} = -10.65 + 0.072\log(BlockSize) + 1.0007 \log(TransactionHistory) + 0.222\log(HashRate)$.

Which can be interpreted as a one percent increase in Blocksize resulting in a 0.072 percent increase in Ether price. The same interpretation applies to the other predictors with there respective coefficients. 




```{r augment}
augmented_reg_obj <- reg_obj %>%
             augment()

some_plot <- ggplot(data = augmented_reg_obj, aes(log.price., .hat)) +
  geom_point(color = "purple", size = 1) +
  labs(x="log(price)", y="yhat") +
  scale_y_log10() +
  ggtitle("Y hat over Y")

some_plot
```
The preceding plot points to the non-linear nature of the data set. A regression model with a polynomial term could lead to future improvements.





```{r}
tidy_conf_it <- tidy(reg_obj, conf.int = TRUE)

ggplot(tidy_conf_it, aes(term, estimate, color=term)) +
  geom_point() +
  geom_errorbar(aes(ymin=estimate - 1.96*std.error, ymax=estimate + 1.96*std.error)) +
  ggtitle("Coefficients with standard errrors")
  
  
  
 
  
  

```
The previous graph is misleading due to the scale with a scale adjustment it is more insightful.






```{r}
some_plot_2 <- ggplot(data = augmented_reg_obj, aes(.resid)) + 
  geom_histogram(binwidth = .3, fill = "purple") + 
  labs(x="Residuals", y="Frequency") +
  ggtitle("Distribution of Residuals")
  
  

some_plot_2
```
The majority of the residuals are distributed between -1 and 1 peaking at zero. 



```{r, warning= FALSE}
some_plot_3 <- ggplot(augmented_reg_obj, aes(log.price., .resid)) +
  geom_point(color = "purple", size = 1) +
  geom_hline(color = "yellow", yintercept = 0) +
  labs(x="Y", y="Residuals") +
  ggtitle("Residuals over Ether Price")

some_plot_3 
  



```

The Residuals plotted against Ether Price suggest there is possible non-normality in the distribution of the errors.

##KNN Regression







```{r knn regression, eval=FALSE}
library(caret)

set.seed(20)

train_index <- createDataPartition(ether_df$price, p = .8, list = FALSE, times = 1)

head(train_index)

ether_df_train <- ether_df[train_index, ]

ether_df_test <- ether_df[-train_index, ]

tr_control <- trainControl(method = "repeatedcv",
             number = 10,
             repeats = 3
             )
knn_model <-  train(price ~., data = ether_df,
                    method = "knn",
                    preProcess = c("center", "scale"), 
                    trControl = tr_control, 
                    metric = "RMSE",
                    tuneLength = 10
                    )

knn_model

knn_model$results

y_hat <- predict(knn_model, newdata = ether_df)

predicted <- predict(knn_model, ether_df)

plot(ether_df$price, predicted)
sqrt(sum((predicted - ether_df$price) ^ 2) / length(ether_df)) 

```

